STAT 450 Project: Real Estate

### 2020-03-03 doc/RealEstate_Yuetong_Liu.Rmd
Individual Report

Our client is a real estate data analyst in Metro Vancouver who analyzes the market trend to make a market prediction for his company and future client. He is interested in accurately predicting future mill rates in Metro Vancouver. There are three major sections of the project: Data cleaning, exploratory data analysis, and linear modelling. Data cleaning aggregated our data into summary statistics and also reduced the number of data entries. Exploratory data analysis analyzed correlation between mill rate and othee features in assessment data frame. Linear modelling gave the best goodness of fit and prediction power among reduced linear regression models, Ridge, Lasso and Elastic net.

@@

The dataset was filter based on below conditions:

@@

Mill Rate and 4 features relevent to Mill Rate are selected:

@@

Based on client information, all properties in the same region, classcode, and year should have a unique class rate. Therefore, all data were grouped by **region + classcode + year**. The assessement value and count of each properties within a group was summed up. 
Moreover, we decided to calculate the growth rate of TotalAssessedValue, LandAssessedValue, ImprovementAssessedValue, Budget and Tax Rate for each municipality and their corresponding tax class.

@@

Before we made any prediction on the future mill rate of Metro Vancouver’s real estate market, we did some exploratory analysis to visualize the main characteristics of our dataset and found relationships between mill rate, assessment and municipal budget.

@@

**Goodness of fit** is defined as the extent to which the sample data are consistent with the model, which examines how well the model explains the data. R squared and adjusted R squared are the most well known measures of goodness of fit and higher R squared and adjusted R squares indicate better goodness of fit.

**Prediction power** measures how well models can predict the future values. In this study, we use mean squared prediction power to compare the prediction performance across all the models.  To do that, we will divide the data set into training data, used as building models, and testing data, used as computing mean squared prediction power.

@@

### Selection of variables

A list of significant variables are included in the result of the full model. In the next step, we will refit the model with only the significant variables. Goodness of fit will decline with fewer features, however, the performance of prediction power might be improved.

@@

We build the three models using all features and compare their goodness of fit, and then will examine their prediction power on testing data.

### Neural Network

In the current report, we are mainly focusing on predicting the mill rate using Linear Model. In the future, we will discuss with our supervisor and try to fit a neural network model if possible (R package: neuralnet or keras). Neural Network has the ability to learn and model non-linear and complex relationships. It is flexible and can be used for both classification and prediction. Since we have a large number of data, it makes Neural Network a very good candidate for our prediction model.

Report on the performance of the Neural Network will be included in the next report.
\pagebreak

@@

multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  numPlots = length(plots)
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }
 if (numPlots==1) {
    print(plots[[1]])
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

@@

# Note: tax_pct doesn't have Maple Ridge Rural. We decide to treat Maple Ridge Rural and Maple Ridge with the same tax.

@@

# tax_final is the budget
tax_modified <- tax_final  %>% mutate(X2019=tax_final$X2018*(1+(((tax_final$X2017-tax_final$X2016)/tax_final$X2016 + (tax_final$X2018-tax_final$X2017)/tax_final$X2017)/2)))

# check match between budget and assessTotal*mill_rate
# get predicted tax income for all tax classes and each municipality
assessment_aggregate[,10] <- assessment_aggregate$assessTotal*assessment_aggregate$rate/1000
names(assessment_aggregate)[10] <- paste("predTaxIncome")

dat_pred_income <- assessment_aggregate %>% select(Year,AddressAssessorMunicipalityDesc,TaxClassCode,predTaxIncome)
dat_pred_income <- aggregate(dat_pred_income$predTaxIncome,by=list(Year=dat_pred_income$Year,Municipalities=dat_pred_income$AddressAssessorMunicipalityDesc),FUN=sum)
names(dat_pred_income)[3] <- paste("TaxIncome")

@@

income_budget_compare <- merge(dat_true_income,dat_pred_income,by=c("Municipalities","Year"))
income_budget_compare <- gather(income_budget_compare, TaxType, Amount, Budget:TaxIncome, factor_key=TRUE)

compare_plot1 <- income_budget_compare %>% filter(Municipalities=="Vancouver") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("Figure 1: Budget and Tax Income v.s. Year of Vancouver") 
compare_plot3 <- income_budget_compare %>% filter(Municipalities=="Burnaby") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("Figure 2: Budget and Tax Income v.s Year of Burnaby")
compare_plot1
compare_plot3
```

@@

# correlation between mill rate and budget for each municipality
rate_tax_correlation <- assessment_aggregate %>%
  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
  summarise(cor(rate, tax))
#rate_tax_correlation
# correlation between mill rate and assessment
rate_assess_correlation2 <- assessment_aggregate %>%
  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
  summarise(cor(assessTotal, rate))
#rate_assess_correlation2
names(assessment_aggregate)[1] <- paste("Municipality")
## Correlation for continuous variables
assess_class <- assessment_aggregate %>% ggplot(aes(x=assessTotal,y=rate,group=Municipality,color=Municipality)) + geom_point() + ggtitle("Figure 3: Correlations between Mill Rates and Total Assessment Values of all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

@@

```{r anova, echo=FALSE, warning=FALSE, message=FALSE}
## Correlation for categorical variables
TaxClassCode <- aov(rate ~ factor(TaxClassCode), data = assessment_aggregate)
#summary(TaxClassCode)

@@

```{r full, echo=FALSE, warning=FALSE, message=FALSE}
linear_full<-lm(rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=assessment_aggregate)
# summary(linear_full)

@@

linear_full_fit<-augment(linear_full)
mse_full <- sqrt(sum((linear_full_fit$.resid)^2)/nrow(assessment_aggregate)) 
```
```{r reduced, echo=FALSE, warning=FALSE, message=FALSE}
reduced<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(Municipality)+assessTotal+landTotal, data=assessment_aggregate)
#summary(reduced)

@@

```{r Ridge, echo=FALSE, warning=FALSE, message=FALSE}
library(glmnet)
library(dummies)
dummy_year<-dummy(assessment_aggregate$Year)
dummy_municipal<-dummy(assessment_aggregate$Municipality)
dummy_taxclass<-dummy(assessment_aggregate$TaxClassCode)
# build x matrix
x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_aggregate$assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$tax)

@@

```{r LASSO, echo=FALSE, warning=FALSE, message=FALSE}
# Setting alpha = 1 implements lasso regression
set.seed(450)
lasso_reg <- cv.glmnet(x, y, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

@@

```{r elastic_net, echo=FALSE, warning=FALSE, message=FALSE}
library(caret)
#tibble::as_tibble(assessment_aggregate)
cv_10 = trainControl(method = "cv", number = 10)
elastic_net = train(
  rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount, data = assessment_aggregate,
  method = "glmnet",
  trControl = cv_10
)
```

```{r Prediction power, echo=FALSE, warning=FALSE, message=FALSE}
set.seed(450)
train_ind<-sample(218,218-50)
train<-assessment_aggregate[train_ind,]
test<-assessment_aggregate[-train_ind,]

# Full linear model
newx<-test[,-c(8,9)]
y<-test[,c(8)]
linear_1<-lm(rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=train)
resid<-predict(linear_1,newdata = newx) - y
full_mspe <- sqrt(sum(resid^2)/nrow(test))

# Reduced model
linear_2<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(Municipality)+assessTotal+landTotal,data=train)
resid<-predict(linear_2,newdata = newx) - y
reduced_mspe <- sqrt(sum(resid^2)/nrow(test))

@@

# Elastic Net
#tibble::as_tibble(assessment_aggregate[train_ind,])
cv_10 = trainControl(method = "cv", number = 10)
elastic_net = train(
 rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax, 
 data = assessment_aggregate[train_ind,],
  method = "glmnet",
  trControl = cv_10
)

@@

## Conclusions
Among all five models, the full linear regression model is the most ideal prediction model since it make accurate prediction without losing information. There are some aspects that we could improve/implemented in the future:

@@

- **Outlier**: In most correlation scatter plots, Vancouver always is an outlier. Since it makes up large portion of the market, we would like to further investigate Vancouver as a special case. We are also interested in the effect of different tax classes in predicting mill rate, since the P-value of ANOVA between tax class and mill rate was extremely small.

@@

### Missing Value:

There are 1801 missing values in mill rate (TaxClassTaxRate). Since this is the predictor, we decided to impute these missing values
Based on client information, all properties in the same region, classcode, and year should have a unique class rate

@@

#### In some groups, all entries' mill rates are missing. Entries in these groups are removed. Here is the list of the groups:

>>>>

Our client is a real estate data analyst who is interested in accurately predicting future mill rates in Metro Vancouver. There are three major sections of the project: Data cleaning, exploratory data analysis, and linear modelling. Data cleaning aggregated our data into summary statistics and also reduced the number of data entries. Exploratory data analysis analyzed correlation between mill rate and other features in assessment data frame. Linear modelling gave the best goodness of fit and prediction power among reduced linear regression models, Ridge, Lasso and Elastic net.

@@

The dataset was filtered based on below conditions:

@@

Mill Rate and 4 features relevent to Mill Rate were selected:

@@

Based on client information, all properties in the same region, classcode, and year should have a unique class rate. Therefore, all data were grouped by **region + classcode + year**. The assessement value and count of each properties within a group were summed up. Moreover, we decided to calculate the growth rate of TotalAssessedValue, LandAssessedValue, ImprovementAssessedValue, Budget and Tax Rate for each municipality and their corresponding tax class.

@@

Before making any prediction on the future mill rate of Metro Vancouver’s real estate market, we did some exploratory analysis to visualize the main characteristics of our dataset and help with model selection. Here are two of our major approach:

@@

@@

@@

### Goodness of fit

It is defined as the extent to which the sample data are consistent with the model, which examines how well the model explains the data. R squared and adjusted R squared are the most well known measures of goodness of fit and higher R squared and adjusted R squares indicate better goodness of fit.

### Prediction power

It measures how well models can predict the future values. In this study, we use mean squared prediction power to compare the prediction performance across all the models.  To do that, we will divide the data set into training data, used as building models, and testing data, used as computing mean squared prediction power.

@@

multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {

  library(grid)
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  numPlots = length(plots)
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }
 if (numPlots==1) {

    print(plots[[1]])

  } else {

    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

@@

# Note: tax_pct doesn't have Maple Ridge Rural. We decide to treat Maple Ridge 
# Rural and Maple Ridge with the same tax.

@@

# tax_final is the budget
tax_modified <- tax_final  %>% 
  mutate(X2019=tax_final$X2018*
           (1+(((tax_final$X2017-tax_final$X2016)/tax_final$X2016 +
                  (tax_final$X2018-tax_final$X2017)/tax_final$X2017)/2)))

# check match between budget and assessTotal*mill_rate

```{r}
# get predicted tax income for all tax classes and each municipality
assessment_aggregate[,10] <- 
  assessment_aggregate$assessTotal*assessment_aggregate$rate/1000
```

```{r graphr, echo= FALSE, warning=FALSE, message=FALSE}
names(assessment_aggregate)[10] <- paste("predTaxIncome")

dat_pred_income <- assessment_aggregate %>%
  select(Year,AddressAssessorMunicipalityDesc,TaxClassCode,predTaxIncome)
dat_pred_income <- 
  aggregate(dat_pred_income$predTaxIncome,
            by=list(Year=dat_pred_income$Year,
                    Municipalities=
                      dat_pred_income$AddressAssessorMunicipalityDesc),FUN=sum)
names(dat_pred_income)[3] <- paste("TaxIncome")

@@

income_budget_compare <- merge(dat_true_income,dat_pred_income,
                               by=c("Municipalities","Year"))
income_budget_compare <- gather(income_budget_compare, TaxType, Amount, 
                                Budget:TaxIncome, factor_key=TRUE)

compare_plot1 <- income_budget_compare %>% 
  filter(Municipalities=="Vancouver") %>% 
  ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + 
  geom_line() + 
  ggtitle("Figure 1: Budget and Tax Income v.s. Year of Vancouver")

compare_plot3 <- income_budget_compare %>% 
  filter(Municipalities=="Burnaby") %>%
  ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + 
  geom_line() + 
  ggtitle("Figure 2: Budget and Tax Income v.s Year of Burnaby")
compare_plot1
compare_plot3
```

@@

#rate_assess_correlation2

names(assessment_aggregate)[1] <- paste("Municipality")
## Correlation for continuous variables
assess_class <- assessment_aggregate %>% 
  ggplot(aes(x=assessTotal,y=rate,group=Municipality,color=Municipality)) + 
  geom_point() + 
  ggtitle("Figure 3: Correlations between Mill Rates and Total Assessment Values of all municipalities") + 
  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black") + 
  theme(plot.background = element_rect(fill = "darkgrey"), legend.key.size = unit(0.2, "cm"))

@@

```{r anova, echo=TRUE, warning=FALSE, message=FALSE}
## Correlation for categorical variables
TaxClassCode <- aov(rate ~ factor(TaxClassCode), data = assessment_aggregate)
#summary(TaxClassCode)

@@

```{r full, echo=TRUE, warning=FALSE, message=FALSE}
## Full Linear model
linear_full<-lm(rate~factor(Municipality)+factor(Year)+
                  factor(TaxClassCode)+assessTotal+landTotal+
                  improvementTotal+propertyCount+tax,data=assessment_aggregate)
# summary(linear_full)

@@

linear_full_fit<-augment(linear_full)
mse_full <- sqrt(sum((linear_full_fit$.resid)^2)/nrow(assessment_aggregate)) 
```

```{r reduced, echo=TRUE, warning=FALSE, message=FALSE}
## Reduced Linear model
reduced<-lm(rate~factor(Year)+factor(TaxClassCode)+
              factor(Municipality)+assessTotal+landTotal, data=assessment_aggregate)
#summary(reduced)

@@

```{r Ridge, echo=TRUE, warning=FALSE, message=FALSE}
## Ridge
library(glmnet)
library(dummies)
dummy_year<-dummy(assessment_aggregate$Year)
dummy_municipal<-dummy(assessment_aggregate$Municipality)
dummy_taxclass<-dummy(assessment_aggregate$TaxClassCode)
# build x matrix
x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_aggregate$assessTotal,
         assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,
         assessment_aggregate$propertyCount,assessment_aggregate$tax)

@@

```{r LASSO, echo=TRUE, warning=FALSE, message=FALSE}
# LASSO
# Setting alpha = 1 implements lasso regression
set.seed(450)
lasso_reg <- cv.glmnet(x, y, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

@@

```{r elastic_net, echo=TRUE, warning=FALSE, message=FALSE}
# Elastic Net
library(caret)
#tibble::as_tibble(assessment_aggregate)
cv_10 = trainControl(method = "cv", number = 10)
elastic_net = train(
  rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+
    landTotal+improvementTotal+propertyCount, data = assessment_aggregate,
  method = "glmnet",
  trControl = cv_10
)
```

```{r Prediction power, echo=TRUE, warning=FALSE, message=FALSE}
# Prediction power
set.seed(450)
train_ind<-sample(218,218-50)
train<-assessment_aggregate[train_ind,]
test<-assessment_aggregate[-train_ind,]

# Full linear model
newx<-test[,-c(8,9)]
y<-test[,c(8)]
linear_1<-lm(rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+
               assessTotal+landTotal+improvementTotal+propertyCount+tax,data=train)
resid<-predict(linear_1,newdata = newx) - y
full_mspe <- sqrt(sum(resid^2)/nrow(test))

# Reduced model
linear_2<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(Municipality)+
               assessTotal+landTotal,data=train)
resid<-predict(linear_2,newdata = newx) - y
reduced_mspe <- sqrt(sum(resid^2)/nrow(test))

@@

# Elastic Net
#tibble::as_tibble(assessment_aggregate[train_ind,])
cv_10 = trainControl(method = "cv", number = 10)
elastic_net = train(
 rate~factor(Municipality)+factor(Year)+factor(TaxClassCode)+assessTotal+
   landTotal+improvementTotal+propertyCount+tax, 
 data = assessment_aggregate[train_ind,],
  method = "glmnet",
  trControl = cv_10
)

@@

## Conclusions

Among all five models, the full linear regression model is the most ideal prediction model since it make accurate prediction without losing much information. There are some aspects that we could improve/implemented in the future:

@@

- **Outlier**: In most correlation scatter plots, Vancouver always is an outlier. Since it makes up large portion of the market, we would like to further investigate Vancouver as a special case. Moreover, we are also interested in the effect of different tax classes in predicting mill rate, since the P-value of ANOVA between tax class and mill rate was extremely small.

@@

### Missing Value and Imputation:

Of 700,000 data entries, there are 1801 missing values in mill rate. Since this is the predictor, we decided to impute these missing values.

Based on client information, all properties in the same region, classcode, and year should have a unique class rate. Therefore, we followed the below procedure to impute mill rate.

@@

#### No mill rate in whole group:

All entries' mill rates in some groups are missing; therefore all entries in these groups were removed. Here is the list of the groups:
